---
title: "Untitled"
author: "Cindy Hou"
date: "8/21/2021"
output: html_document
---

Part 1: Import libraries & data

```{r}
library(car)
library(MASS)
library(leaps)
library("randomForest")

train <- read.csv("~/Desktop/Kaggle/titanic/train.csv", stringsAsFactors = TRUE)
test <- read.csv("~/Desktop/Kaggle/titanic/test.csv", stringsAsFactors=TRUE)
test$Survived <- NA

#combining both sets of data for later cleaning
titanicAll <- rbind(train, test)
```

Familiarizing myself with the data
```{r}
head(train)
head(test)

str(train)
str(test)
str(titanicAll)
```

2. Data cleaning & dealing with missing values
```{r}
# Checking for missing values
colSums(titanicAll == "")
colSums(is.na(titanicAll))

table(titanicAll$Embarked)
```
There are only 2 values missing in embarked, so I'm filling the missing values in with the most frequent Embarked value by far, "S." Similarly, since there are is only 1 missing fare value, I'm filling it in with the mean fare value.

```{r}
titanicAll$Embarked[titanicAll$Embarked == ""] <- "S"

titanicAll$Fare[is.na(titanicAll$Fare)] = mean(titanicAll$Fare, na.rm=TRUE)
```

As there are so many Cabin values missing (> half the entries), I'm choosing not to take the Cabin values into account at all. 
```{r}
titanicAll <- subset(titanicAll, select = -Cabin)
```

Lastly, there are a decent amount of missing ages, so I will be more careful in filling in the values as simply filling it in with the mean can have a large effect on the final predictive model. Instead, I'm using linear regression to predict age based on known values of PClass, sex, fare, and embarked.
```{r}
lm.age <- lm(Age ~ Fare + Pclass + Sex + Embarked, data = titanicAll)

#extract PClass, Sex, Fare, and Embarked values from empty rows
extractFromMissingAge <- titanicAll[is.na(titanicAll$Age), c("Pclass", "Sex", "Fare", "Embarked")]

newAge <- predict(lm.age, newdata = extractFromMissingAge)

titanicAll$Age[is.na(titanicAll$Age)] = newAge

# making sure there are no more missing values
colSums(titanicAll == "")
colSums(is.na(titanicAll))
```

After cleaning the data, I'm setting factors and separating the data back into the original train and test sets
```{r message=FALSE, warnings=FALSE}
titanicAll$Pclass <- as.factor(titanicAll$Pclass)
titanicAll$Sex <- as.factor(titanicAll$Sex)
titanicAll$Embarked <- as.factor(titanicAll$Embarked)

trainNew <- subset(titanicAll, PassengerId == train$PassengerId)

trainNew$Survived <- as.factor(trainNew$Survived)

testNew <- subset(titanicAll, PassengerId != train$PassengerId)


```

4. Model 2: Model using random forest
```{r}
rfModel <- randomForest(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked, data = trainNew, ntree = 500, mtry = 3, nodesize = 0.01*nrow(test), importance=TRUE)

predictedSurvivalTest <- predict(rfModel, newdata = testNew)

testNew$Survived <- predictedSurvivalTest
```

```{r}
# preparing submission
submission <- data.frame(testNew$PassengerId, testNew$Survived)
colnames(submission) <- c("PassengerId", "Survived")

write.csv(submission,"/Users/indiachina/Desktop/Kaggle/titanic/titanic_prediction.csv",row.names = FALSE)
```

